Mendelian randomization for causal inference in Genome-wide association studies

### 2022-02-07 mr-review.Rmd
text following MRAID paper

We discuss PMR-Egger below in the context of methods that deal with horizontal pleiotropy.

@@

@@

In the case of continuous exposures and outcomes, we extend the methods and models for 
single-SNP MR to accommodate multiple SNPs. For two stage analyses, we perform 
a sequence of two linear regressions, as in the single-SNP MR models (Equations \@ref(eq:lm3) and \@ref(eq:lm4)). 
For an analysis
with $S$ SNP instruments, one performs two stage analysis by first regressing the 
exposure $n$ vector $X$ on the $n$ by $S$ matrix of SNP instruments. Taking the fitted exposure values 
from this regression, the second regression involves regressing the outcome variable $Y$ on 
the fitted values.

@@

Horizontal pleiotropy refers to the presence of multiple pathways from a genetic 
variant to the outcome. One of these multiple pathways may contain the exposure variable.
The fact that there are multiple pathways from the genetic variant to the outcome violates
the MR assumptions. @verbanck_detection_2018 documented widespread horizontal pleiotropy 
in GWAS. This finding motivated the development of MR-PRESSO and inspired 
further research into methods to model and to account for horizontal pleiotropy [@verbanck_detection_2018].

>>>>

We discuss PMR-Egger below in the context of methods that model horizontal pleiotropy.

@@

Choosing a subset of the available SNPs to serve as instruments is one consideration that arises 
in MR studies. The solution to this task differs among published methods. Early methods chose a 
set of independent SNPs. However, this approach can be problematic when there are multiple causal 
SNPs in LD with each other. In this setting, discarding some of the causal SNPs may only capture a small
proportion of trait variance explained by the exposure and lead to a power loss in MR [@burgess_review_2017;@burgess_mendelian_2013;@burgess_combining_2016;@burgess_bias_2011]. In support of
this is the finding that including multiple correlated SNPs in transcriptome-wide association
studies improves power compared to using independent SNPs only [@liu2021multi;@gusev_integrative_2016;@yuan_testing_2020;@zeng_non-parametric_2017].

@@

In the case of continuous exposures and outcomes, researchers have taken several approaches to extend the methods and models for 
single-SNP MR to accommodate multiple SNPs. For two stage analyses, one may perform 
a sequence of two linear regressions, as in the single-SNP MR models (Equations \@ref(eq:lm3) and \@ref(eq:lm4)). 
For an analysis
with $S$ SNP instruments, one performs two stage analysis by first regressing the 
exposure $n$ vector $X$ on the $n$ by $S$ matrix of SNP instruments. Taking the fitted exposure values 
from this regression, the second regression involves regressing the outcome variable $Y$ on 
the fitted values.

@@

Horizontal pleiotropy refers to the presence of multiple pathways from a genetic 
variant to the outcome. One of these multiple pathways may contain the exposure variable.
The fact that there are multiple pathways from the genetic variant to the outcome violates
standard MR assumptions. @verbanck_detection_2018 documented widespread horizontal pleiotropy 
in GWAS. This finding motivated the development of MR-PRESSO and inspired 
further research into methods to model and to account for horizontal pleiotropy [@verbanck_detection_2018].

Early MR analyses avoided confounding from horizontal pleiotropy by discarding 
instrumental SNPs that might be associated with the outcome [@verbanck_detection_2018;@zhu_causal_2018;@zeng_causal_2019-1;@zeng_causal_2019-2].
More recent methods have attempted to model horizontal pleiotropy [@morrison_mendelian_2020;@yuan_testing_2020;@yuan_likelihood_2021]. CAUSE [@morrison_mendelian_2020] mdoels two varieties of horizontal pleiotropy:
